### 2018-05-30 Physics Notes/Paper 2/Paper 2.tex
Physics paper 2

\documentclass[12pt]{article}
\usepackage{../../format}
\lhead{A Level Physics}
\usepackage{enumitem}
\setcounter{secnumdepth}{4}
\usepackage{float}

@@

$ $\\

@@

\includegraphics[width=6cm]{foil.jpg}  
    \end{minipage}\hfill
    \begin{minipage}{0.45\textwidth}
        \centering
\begin{tabularx}{\textwidth}{|X|X|}
\hline
Observation&Explanation\\
\hline
Most electrons pass all the way through&Atoms are mostly empty space\\
\hline
Some are deflected&The atom has a positive centre\\
\hline
Some are deflected by significant angles&The positive charge is condensed in a small area\\
\hline
\end{tabularx}
    \end{minipage}
\end{figure}
\subsubsection{$\alpha$, $\beta$ and $\gamma$ radiation}
\begin{tabularx}{\textwidth}{|c|X|X|X|}
\hline
&Alpha&Beta&Gamma\\
\hline
Nature&2 Protons+2 Neutrons&High speed electron or positron&High energy photon\\
\hline
Range&Up to 10cm&Up to 1m&Infinite\\
\hline 
Deflection in a magnetic field&Deflected&Opposite direction to $\alpha$ particles and more easily deflected&Not deflected\\
\hline
Absorption&Paper&Aluminium&Lead \\
\hline
Ionisation&$10^4$ ions per mm&100 ions per mm&Very weak ionising effect\\
\hline
Energy of each particle&Constant for a given source&Varies up to a maximum for a given source&Constant for a given source\\
\hline
\end{tabularx}
\paragraph{Inverse square law for gamma radiation}
$$I=\frac{k}{x^2}$$
\paragraph{Safety}$ $\\
Those working in areas of high radiation, for example nuclear power plants, should wear film badges which monitor a persons exposure to radiation
\paragraph{Background radiation}$ $\\
Background radiation has many sources:
\begin{itemize}
\item Air (Radon gas)
\item Medical
\item Ground and buildings
\item Food and drink
\end{itemize}
\paragraph{Uses of radiation in medicine}$ $\\
The health benefits from using radiation in medicine almost always outweigh the risks, but doctors must protect themselves as repeated exposure is dangerous
\end{document}

>>>>

\documentclass[12pt]{article}
\usepackage{../../format}
\lhead{A Level Physics}
\usepackage{enumitem}
\setcounter{secnumdepth}{6}
\usepackage{float}
\usepackage[english]{babel}

\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{0ex}%
    {-3.25ex plus -1ex minus -0.2ex}%
    {1.5ex plus 0.2ex}%
    {\normalfont\normalsize\bfseries}}
\makeatother

\makeatletter
\renewcommand{\subparagraph}{\@startsection{subparagraph}{4}{0ex}%
    {-3.25ex plus -1ex minus -0.2ex}%
    {1.5ex plus 0.2ex}%
    {\normalfont\normalsize\bfseries}}
\makeatother

@@

@@

\includegraphics[width=6cm]{foil.jpg}  
    \end{minipage}\hfill
Observation&Explanation\\
Most electrons pass all the way through&Atoms are mostly empty space\\
Some are deflected&The atom has a positive centre\\
Some are deflected by significant angles&The positive charge is condensed in a small area\\
    \end{minipage}
\end{figure}
\subsubsection{$\alpha$, $\beta$ and $\gamma$ radiation}
\begin{tabularx}{\textwidth}{|c|X|X|X|}
&Alpha&Beta&Gamma\\
Nature&2 Protons+2 Neutrons&High speed electron or positron&High energy photon\\
Range&Up to 10cm&Up to 1m&Infinite\\
Deflection in a magnetic field&Deflected&Opposite direction to $\alpha$ particles and more easily deflected&Not deflected\\
Absorption&Paper&Aluminium&Lead \\
Ionisation&$10^4$ ions per mm&100 ions per mm&Very weak ionising effect\\
Energy of each particle&Constant for a given source&Varies up to a maximum for a given source&Constant for a given source\\
\paragraph{Inverse square law for gamma radiation}
$$I=\frac{k}{x^2}$$
\paragraph{Safety}
Those working in areas of high radiation, for example nuclear power plants, should wear film badges which monitor a persons exposure to radiation
\paragraph{Background radiation}
Background radiation has many sources:
\item Air (Radon gas)
\item Medical
\item Ground and buildings
\item Food and drink
\paragraph{Uses of radiation in medicine}
The health benefits from using radiation in medicine almost always outweigh the risks, but doctors must protect themselves as repeated exposure is dangerous
\subsubsection{Radioactive decay}
$$\frac{\Delta N}{\Delta t}=-\lambda N$$
$$N=N_0e^{-\lambda t}$$
$$A=\lambda N$$
$$A=A_0e^{-\lambda t}$$
$$m=m_0e^{-\lambda t}$$
$$C=C_0e^{-\lambda t}$$
$$T_{\frac{1}{2}}=\frac{\ln 2}{\lambda}$$
$N$ - Number of nuclei\\
$\lambda$ - Decay constant\\
$A$ - Activity\\
$m$ - Mass of the isotope\\
$C$  - Count rate on a Geiger counter\\
$T_{\frac{1}{2}}$ - Half life of the isotope
\paragraph{Constant decay probability}
There is a constant decay probability of an isotope, and the decays happen at random, this means that the number of nuclei that decay in a certain time interval depends only on the total number of nuclei present.
\paragraph{Applications}
\subparagraph{Carbon dating}
The half life of carbon-14 is a known value and so can be used to determine the age of objects. By looking at the activity of a live equivalent of equal mass and comparing it, the age can be found.\\
\textbf{Example}:\\
Activity =0.28Bq, Live activity=1.3Bq, half life = 5570 years
$$T_{\frac{1}{2}}=1.76\times 10^{11}$$
$$\lambda=\frac{\ln 2}{1.76 \times 10^{11}}=3.95\times 10^{12}$$
$$A=A_0e^{-\lambda t} \Rightarrow 0.28=1.3e^{-\lambda t} \Rightarrow \lambda t=1.535$$
$$t=\frac{1.535}{\lambda}=\frac{1.535}{3.95\times 10^{-12}}=3.88\times 10^{11}s$$
\subparagraph{Radioactive tracers}
Radioactive tracers should:
\item Have a half life long enough for the necessary measurements to be made, but short enough to decay quickly after use
\item Emit $\beta$ or $\gamma$ radiation so that it can be detected outside the body
\subsubsection{Nuclear instability}
\includegraphics[width=8cm]{nz_plot.png}
\includegraphics[width=6cm]{nz_changes.png}
Electron capture would be a horizontal line to the left
\paragraph{Nuclear excited states}
The nuclear excited state is where the nucleus has more energy than its basic state, to counteract this, the nucleus gives off the energy in the form of a $\gamma$ ray\\
This can be used to make artificial gamma ray emitters
\subsubsection{Nuclear radius}
\paragraph{Closest approach of alpha particles}
By determining the percentage of $\alpha$ particles that are deflected by more than 90\degree, around 1 in 10,000. Given n layers of atoms, the probability is 1 in 10,000n. Then determine the cross sectional area of the radius of the atom and nucleus as a ratio. 
$$\frac{\frac{1}{4}\pi d^2}{\frac{1}{4}\pi D^2}=\frac{1}{10,000n}$$
$$d^2=\frac{D^2}{10000n}$$
A typical value of n is 10,000
$$d=\frac{D}{10000}$$
As the diameter of an atom can be more easily measured, this gives a result of around 1fm
\paragraph{Electron diffraction}
The diffraction of electrons can be used to give a more accurate answer for the nuclear radius. Electrons are accelerated so that their de Broglie wavelength is approximately the radius of the nucleus so that they diffract best.\\
There is an angle $\theta_{min}$ away from the normal at which there is a local minimum intensity, using the relation
$$R\sin\theta_{min}=0.61\lambda$$
The value of R can be found, $\lambda$ is the wavelength of the electrons
\end{document}
